---
title: "Data Project Part 1"
author: Carri Beshears, Samantha Sears, Samuel Johnson, Catalina Perez, and Phuong
  Tseng
date: "2022/2/12"
output:
  pdf_document: default
  html_document: default
header-includes: \usepackage{epsfig,graphicx,graphics,amssymb,amsmath,latexsym,amsfonts,url,
  lscape, multirow, color}
---
***Data Project Part 1***

Research Question: Do women in CA have <b>higher rates</b> of Chlamydia than men? 
You can find our analysis on Github: [https://github.com/phtseng/PBHLTH_142_Group/blob/master/Data_Project_Part_1.Rmd]

1. [2 marks] Identify a statistical test to apply to your data (must be a concept we covered in part III of the course). In plain language, write the question you are trying to answer.

Chlamydia rates continue to climb in the United States each year, according to the CDC. Is there a 
difference of disease prevalence among individuals that identify as women and individuals that identify as men in the 
United States between the years of 2000 and 2014? To start to answer this question, we will conduct a
two-sample t-test for proportions comparing aggregated data from each state. Chlamydia 
cases that were entered with an “unknown” gender will be removed from this analysis due poor clarity of 
Its meaning.  

2. [2 marks] What assumptions are required by the method you chose in 5.1?
Show how you assessed whether these assumptions are met by your dataset.

Independent populations, simple random sampling and a large sample size are required 
for the two sample t-test for proportions statistical  test.

The sampling frame used CDC STI data from 2000-2014 which consists of aggregated data reported 
to the CDC from trustworthy and validated sources. The participants are randomly selected using 
simple random sampling. This method for probability sampling ensures that every man or woman 
within the population has the same probability of being selected for the study. Therefore, the process for choosing a person is not biased. Also, random samples are from independent populations. 
The study sample size is large which ensures the distribution of the proportion is normally distributed.

3. [2 marks] Explain why this test is appropriate for the data you have and the question you are trying
to answer. 

The two sample t-test for proportions is appropriate for this data set because we are comparing the portions of 2 distinct populations within our data set. Our proportions are derived from a large sample size, has an approximately normal distribution, and represents all cases reported in the US during 2000-2014. 

4. [2 marks] Clearly state the null and alternative hypotheses for this test.

H0: There is no difference between the proportion of chlamydia seen in men and women in the United States during 2000-2014

P1 = p2 | p1 - p2 = 0

Ha: There is a difference between the proportion of chlamydia seen in men and women in the United States during 2000-2014

P1 != p2 | p1 - p2 != 0 (two-sided)

5. [2 marks] Include the R code you used to generate your results. Annotate your code to help us follow
your reasoning.

6. [4 marks] Present your results in a clear summary. This should include both a text summary and a
table or figure with appropriate labeling.


7. [4 marks] Interpret your findings. Include a statement about the evidence, your conclusions, and the
generalizability of your findings.
Additional studies should be done. Only two genders included, no inclusion of MSM or FSF. 


#### Set up the packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dpi=300,fig.width=8)

x <-
  c("data.table",
    "readxl",
    "dplyr",
    "broom",
    "ggplot2",
    "usethis",
    "sf",
    "gridExtra",
    "knitr",
    "here")

lapply(x, require, character.only = TRUE)
#tinytex::install_tinytex() #To load tinytext() so we can knit to Word document
```
To load the libraries that we need without repeating the library function, we used the lapply() function in R to load them all at once.


*Data Project PART 3*

#### Import Data
```{r Import Data}
sti_data <- read_xlsx("sti_data.xlsx", sheet = "2000-2014")
```
We use the read_xlsx function to read in the data set and specify the excel sheet then assign it to an object called <b>sti_data</b>.

#### Data Cleaning
```{r Clean Data for Analysis}
df <-
  sti_data %>%
  select(ID, State, Gender, Disease, Count, Population, Rate) %>%
  filter(Gender != "Unknown", Disease=="Chlamydia") %>% 
  mutate(
    Count = as.numeric(Count),
    Population = as.numeric(Population),
    Proportion = Count / Population)

View(df)
```
We use the dplyr library and pipe operator to select the relevant columns then filter our records to remove "unknown" gender then create the count and population.

#### Data Cleaning Continue
```{r Remove Districts from dataset}
# Remove District Because Data Collectors Stated Errors in These States When Sampling then Remove NAN
df2 <-  filter(df, State != "District of Columbia", State != "Alaska", State != "Mississippi") %>% na.omit(df)

# Confirm Numeric (removing this since it's in line 437)
#Data = mutate(Count = as.numeric(Count), Population = as.numeric(Population))

# Create Proportion (see line 37)
#Data = mutate(Data, Proportion = Count/Population)
```

#### Question 5 
[2 marks] Include the R code you used to generate your results. Annotate your code to help us follow your reasoning.

```{r Question 5}
show <- df2 %>% group_by(Disease, Gender) %>% mutate(Rate=as.numeric(Rate)) %>% filter(Disease=='Chlamydia') %>% summarize(TotalCount = sum(Count), TotalPop = sum(Population), TotalRate = sum(Rate)) 

show #table with the totals

dfframe <- data.frame(Gender = c("Female_w_C", "Male_w_C"), count=c(13529616, 4774545), pop=c(2535475282,2464469983), prop=c(13529616 /  2535475282, 4774545 / 2464469983))

dfframe 

prop.test(x=c(13529616, 4774545), n=c(13529616 + 2535475282 , 4774545 + 2464469983)) #p-value < 2.2e-16

```

#### Question 6 
[4 marks] Present your results in a clear summary. This should include both a text summary and a table or figure with appropriate labeling.
```{r Question 6}
Data <-  df2

# Plotting
Plot = ggplot(data = Data, aes(x = Proportion))
Plot = Plot + geom_histogram(aes(y = stat(density)), binwidth = 0.000055,color='black',fill='grey')
Plot = Plot + geom_density(col = "red", size = 2.5)
Plot = Plot + labs(title = "Distribution of Proportion of Chlamydia Cases for Each State")
Plot = Plot + labs(subtitle = "Stratified by Gender (Male or Female)")
Plot = Plot + labs(x = "Sampling Proportion of Cases")
Plot = Plot + labs(y = "Count")
Plot = Plot + facet_wrap(~ Gender,scales = "free")
Plot = Plot + theme_classic()
Plot

# QQ Plot
QQ_Plot = ggplot(data = Data, aes(sample = Proportion)) 
QQ_Plot = QQ_Plot + stat_qq() + stat_qq_line() 
QQ_Plot = QQ_Plot + facet_wrap(~ Gender)
QQ_Plot = QQ_Plot + theme_classic()
QQ_Plot

# Confidence Intervals
Inference_Data = group_by(Data, Gender)
Summary_Gender = summarize(Inference_Data, Prop = mean(Proportion), Standard = sd(Proportion), Margin_95 = 1.96*Standard/sqrt(n()), Lower = Prop - Margin_95, Upper = Prop + Margin_95)
tibble(Summary_Gender)
```

#### Question 7 
[4 marks] Interpret your findings. Include a statement about the evidence, your conclusions, and the
generalizability of your findings.
Additional studies should be done. Only two genders included, no inclusion of MSM or FSF. 

```{r Question 7}

```



